---
title: "mrubiog2000_PeerCommentary03_aer13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Some of My Best Friends are Zombies...

**Set working directory using setwd() function and load "zombies.csv" dataset into R**

```{r}
setwd("/Users/AbbyRobinson/Desktop/BIOANTHRO STATS")
zombies <- read.csv("zombies.csv", header = TRUE)

```

```{r}
##Miguel: I cannot acces the data this way because I am using a different computer. Next time you can try the following so everyone with whom you share your code can access it:

library(curl)

f <- curl("https://raw.githubusercontent.com/fuzzyatelin/fuzzyatelin.github.io/master/AN588_Fall21/zombies.csv")
zombies <- read.csv(f, header = TRUE, sep = ",", stringsAsFactors = FALSE)
```


##1. Calculate the population mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, and years of education). NOTE: You will not want to use the built in var() and sd() commands as these are for samples.

###height 
  can you just use the mean function to get the population mean??
```{r}
population_mean_height <- mean(zombies$height)
population_mean_height
```

Calculating population standard deviation 
```{r}

x <- zombies$height

pop_v <- function(x) {
    sum((x - mean(x))^2)/(length(x))
}

pop_sd <- function(x) {
    sqrt(pop_v(x))
}
pop_sd(x)
```

###weight 

 mean
```{r}
population_mean_weight <- mean(zombies$weight)
population_mean_weight
```

standard deviation 
```{r}
x <- zombies$weight
pop_sd(x)
```

###age

 mean
```{r}
population_mean_age <- mean(zombies$age)
population_mean_age
```

standard deviation 
```{r}
x <- zombies$age
pop_sd(x)
```

###number of zombies killed

 mean
```{r}
population_mean_zombies_killed <- mean(zombies$zombies_killed)
population_mean_zombies_killed
```

standard deviation 
```{r}
x <- zombies$zombies_killed
pop_sd(x)
```

###years of education 

 mean
```{r}
population_mean_years_of_education <- mean(zombies$years_of_education)
population_mean_years_of_education
```

standard deviation 
```{r}
x <- zombies$years_of_education
pop_sd(x)
```

```{r}
##Miguel: Good job. You can directly apply the function "mean()" or the function you created to the vector to zombies$___. This way you avoid having to create each time an object called "x" that might be confusing as it is the same for the 5 sd of the variables.
```



##2. Use {ggplot} to make boxplots of each of these variables by gender.

```{r}
library(ggplot2)
zom.height <- zombies$height
```

###height 

```{r}
z <- ggplot(data = zombies, aes(x = gender, y = height))
z <- z + geom_boxplot()

z
```

###weight

```{r}
z <- ggplot(data = zombies, aes(x = gender, y = weight))
z <- z + geom_boxplot()

z
```

###age

```{r}
z <- ggplot(data = zombies, aes(x = gender, y = age))
z <- z + geom_boxplot()

z
```

###number of zombies killed

```{r}
z <- ggplot(data = zombies, aes(x = gender, y = zombies_killed))
z <- z + geom_boxplot()

z
```

###years of education

```{r}
z <- ggplot(data = zombies, aes(x = gender, y = years_of_education))
z <- z + geom_boxplot()

z
```

```{r}
##Miguel: Good job. You can draw all plots at once using the package "gridExtra" and the function "grid.arrange("names of the plots", ncol=__)". Again, be careful about giving the same name (z) to every objet you code.
```


##3. Use {ggplot} to make scatterplots of height and weight in relation to age. Do these variables seem to be related? In what way?

###height 
```{r}

z <- ggplot(data = zombies, aes(x = age, y = height, color = factor(gender)))
z <- z + geom_point()
z <- z + theme(legend.position = "bottom", legend.title = element_blank())  
z
```

###weight 
```{r}
z <- ggplot(data = zombies, aes(x = age, y = weight, color = factor(gender)))
z <- z + geom_point()
z <- z + theme(legend.position = "bottom", legend.title = element_blank())  
z
```
both the height and the weight plots show a direct, positive relationship between age and height, and age and weight. I found it surprising that there seems to be a stronger relationship between age and height than there is for age and weight... I would expect there to be a stronger relationship between age and height in people's 20's because humans don't usually grow taller after like 18.... or maybe thats just me... 

```{r}
##Miguel: Awesome plots! I really like how you colored the dots for male and female (I tried to do so but I failed) and added the legend at the bottom of the plot. 
##Miguel: It is not necessary to do several lines of code, you can just write: "z <- ggplot(______) + geom_point()" in one line and it works. Probably the legend can be added to the single line after a "+" too.
```



##4. Using histograms and Q-Q plots, check whether the quantitative variables seem to be drawn from a normal distribution. Which seem to be and which do not (hint: not all are drawn from the normal distribution)? For those that are not normal, can you determine from which common distribution they are drawn?

###height 

```{r}
h <- zombies$height
hist(h) #histogram shows a normal distribution around the mean 
population_mean_height
```

```{r}
qqnorm(h, main = "QQ plot of Zombie Killer Height")
qqline(h, col = "gray") #normally distributed 
```

###weight

```{r}
w <- zombies$weight
hist(w) #histogram shows a normal distribution around the mean 
population_mean_weight
```

```{r}
qqnorm(w, main = "QQ plot of Zombie Killer Weight")
qqline(w, col = "gray") #normally distributed 
```
###age 

```{r}
a <- zombies$age
hist(a) #histogram shows a normal distribution around the mean 
population_mean_age
```

```{r}
qqnorm(a, main = "QQ plot of Zombie Killer Age")
qqline(a, col = "gray") #normally distributed 
```

###number of zombies killed 

```{r}
k <- zombies$zombies_killed
hist(k) #histogram is skewed towards lower numbers 
population_mean_zombies_killed
```

```{r}
qqnorm(k, main = "QQ plot of Zombie Kills")
qqline(k, col = "gray") #WOAH!! super weird! 
```

###years of education 

```{r}
e <- zombies$years_of_education
hist(e) #histogram is skewed towards lower numbers 
population_mean_years_of_education
```

```{r}
qqnorm(e, main = "QQ plot of Zombie Killer Years of Education")
qqline(e, col = "gray") #QQplot looks just like plot zombie kills 
```


```{r}
##Miguel: Good job. You can use the "par()" function to show several plots at the same time in R histograms.
```


##For those that are not normal, can you determine from which common distribution they are drawn? 

number of zombies killed and years of education are not normal because these data are not random. To determine from which common distribution they are drawn, you can look at the data itself...
```{r}
zombies$zombies_killed
zombies$years_of_education
```
these data are independent counts of integer numbers, which would follow the Poisson distribution 


##5. Now use the sample() function to sample ONE subset of 30 zombie survivors (without replacement) from this population and calculate the mean and sample standard deviation for each variable.

###height 
```{r}
?sample()
zombies
sample_height <- sample(zombies$height, size = 30, replace = FALSE, prob = NULL) 
sample_height
mean(sample_height)
sd(sample_height)
```

##Also estimate the standard error for each variable, and construct the 95% confidence interval for each mean. Note that for the variables that are not drawn from the normal distribution, you may need to base your estimate of the CIs on slightly different code than for the normal…

standard error

```{r}
se_height <- (sd(sample_height)/sqrt(30))
se_height
```

construct 95% confidence interval for the mean 

```{r}
x <- sample_height
m <- mean(sample_height)

normalCI = function(x, CIlevel = 0.95) {
    upper = m + qnorm(1 - (1 - CIlevel)/2) * sqrt(var(x)/length(x))
    lower = m + qnorm((1 - CIlevel)/2) * sqrt(var(x)/length(x))
    ci <- c(lower, upper)
    return(ci)
}
normalCI(x, 0.95)
```

###weight 

  create sample  
```{r}
sample_weight <- sample(zombies$weight, size = 30, replace = FALSE, prob = NULL) 
sample_weight
```

  mean 
```{r}
mean(sample_weight)
```
  standard deviation 
```{r}
sd(sample_weight)
```

  standard error 
```{r}
se_weight <- (sd(sample_weight)/sqrt(30))
se_weight
```
  95% Confidence Interval 
```{r}
x <- sample_weight
m <- mean(sample_weight)
normalCI(x, 0.95)
```

###age

  create sample  
```{r}
sample_age <- sample(zombies$age, size = 30, replace = FALSE, prob = NULL) 
sample_age
```

  mean 
```{r}
mean(sample_age)
```
  standard deviation 
```{r}
sd(sample_age)
```

  standard error 
```{r}
se_age <- (sd(sample_age)/sqrt(30))
se_age
```
  95% Confidence Interval 
```{r}
x <- sample_age
m <- mean(sample_age)
normalCI(x, 0.95)
```

###number of zombies killed

  create sample  
```{r}
sample_zombies_killed <- sample(zombies$zombies_killed, size = 30, replace = FALSE, prob = NULL) 
sample_zombies_killed
```
  mean 
```{r}
mean(sample_zombies_killed)
```

  standard deviation 
```{r}
sd(sample_zombies_killed)
```

  standard error 
```{r}
se_zombies_killed <- (sd(sample_zombies_killed)/sqrt(30))
se_zombies_killed
```

95% Confidence Interval 

Write function to calculate confidence intervals under Poisson Distribution 
```{r}
x <- mean(sample_zombies_killed)


poisson95CI = function(x) {
    upper = x + 1.96*(sqrt(x/30))
    lower = x - 1.96*(sqrt(x/30))
    ci <- c(lower, upper)
    return(ci)
}

poisson95CI(x)
```

```{r}
##Miguel: I used 1.96 for the upper, but 2.58 for the lower. I'm not sure way, but those were the numbers I found when seeking for how to do the CI for the Poisson Distribution in google. I think I will ask the professor because I do not understand it well.
```


###years of education 
  create sample  
```{r}
sample_years_of_education <- sample(zombies$years_of_education, size = 30, replace = FALSE, prob = NULL) 
sample_years_of_education
```
  mean 
```{r}
mean(sample_years_of_education)
```

  standard deviation 
```{r}
sd(sample_years_of_education)
```

  standard error 
```{r}
se_years_of_education <- (sd(sample_years_of_education)/sqrt(30))
se_years_of_education
```

95% Confidence Interval 
```{r}
x <- mean(sample_years_of_education)
poisson95CI(x)
```

```{r}
##Miguel: Good job. For this question, I would use the sample function to obtain 30 individual and then calculate the mean and sd of each variable, instead of sampling for each variable (I think is faster), but your way is ok!. To do so, you can use the function "sample_n" from "dplyr" package.
```


##6. Now draw 99 more random samples of 30 zombie apocalypse survivors, and calculate the mean for each variable for each of these samples. Together with the first sample you drew, you now have a set of 100 means for each variable (each based on 30 observations), which constitutes a sampling distribution for each variable. 

###height 
```{r}
set <- NULL
n <- zombies$height
for (i in 1:99) {
    set[i] <- mean(sample(n, size = 30, replace = FALSE, prob = NULL))
}
sampling_distribution <- set[]
sampling_distribution

```

##What are the means and standard deviations of this distribution of means for each variable? How do the standard deviations of means compare to the standard errors estimated in [5]? What do these sampling distributions look like (a graph might help here)? Are they normally distributed? What about for those variables that you concluded were not originally drawn from a normal distribution?

```{r}
mean(sampling_distribution)
sd(sampling_distribution)
```

This standard deviation of the sampling distribution is actually very similar to the standard error from [5] - 0.7010483, COOL! 

```{r}
hist(sampling_distribution) #looks like a normal distrubution 
```
###weight 
```{r}
set <- NULL
n <- zombies$weight
for (i in 1:99) {
    set[i] <- mean(sample(n, size = 30, replace = FALSE, prob = NULL))
}
sampling_distribution <- set[]
sampling_distribution

```

```{r}
mean(sampling_distribution)
sd(sampling_distribution)
```
this is similar to the standard error value fron 5 (3.855582)

```{r}
hist(sampling_distribution) #looks like a normal distribution 
```

###age
```{r}
set <- NULL
n <- zombies$age
for (i in 1:99) {
    set[i] <- mean(sample(n, size = 30, replace = FALSE, prob = NULL))
}
sampling_distribution <- set[]
sampling_distribution

```

```{r}
mean(sampling_distribution)
sd(sampling_distribution)
```
this is similar to the standard error value from 5 (0.4423582)

```{r}
hist(sampling_distribution) #looks like a normal distribution 
```

###zombies killed

```{r}
set <- NULL
n <- zombies$zombies_killed
for (i in 1:99) {
    set[i] <- mean(sample(n, size = 30, replace = FALSE, prob = NULL))
}
sampling_distribution <- set[]
sampling_distribution

```

```{r}
mean(sampling_distribution)
sd(sampling_distribution)
```
this is similar to the standard error value from 5 (0.3571385)

```{r}
hist(sampling_distribution) #looks like a normal distribution 
```

###years of education 
```{r}
set <- NULL
n <- zombies$years_of_education
for (i in 1:99) {
    set[i] <- mean(sample(n, size = 30, replace = FALSE, prob = NULL))
}
sampling_distribution <- set[]
sampling_distribution

```

```{r}
mean(sampling_distribution)
sd(sampling_distribution)
```
this is similar to the standard error value from 5 (0.3956178)

```{r}
hist(sampling_distribution) #looks like a normal distribution 
```

```{r}
##Miguel: Nice job! I learnt a lot from the structure of your loop and how to directly calculate the mean within the loop, thank you! I thought we could not use "Replacement = FALSE" here because there are not enough survivors to sample (99*30), but yours seem to work, so I guess it is ok.
```

```{r}
##Miguel: Really good job! I think you managed to do all questions quite well :) Personally, I sometimes find your coding a bit long due to big number of divisions in the code for each variable, but if this is the way that works for you, it is totally fine.
```


Challenges: 

5. Error in sample.int(length(x), size, replace, prob) : cannot take a sample larger than the population when 'replace = FALSE'

original code: sample(zombies, size = 30, replace = FALSE, prob = NULL), needed to specify zombies$variable 
