---
title: "nkelley_PeerCommentary03_aer13.Rmd"
author: "Abby_Robinson"
date: "10/7/2021"
output: html_document
---

#Some of My Best Friends are Zombies...

**Set working directory using setwd() function and load "zombies.csv" dataset into R**

```{r}
setwd("/Users/AbbyRobinson/Desktop/BIOANTHRO STATS")
zombies <- read.csv("zombies.csv", header = TRUE)
zombies
```
NK: I'm sure this works on your own computer, but anyone who is not on your desktop cannot access the data. 

##1. Calculate the population mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, and years of education). NOTE: You will not want to use the built in var() and sd() commands as these are for samples.

###height 
  can you just use the mean function to get the population mean??
```{r}
population_mean_height <- mean(zombies$height)
population_mean_height
```

Calculating population standard deviation 
```{r}

x <- zombies$height

pop_v <- function(x) {
    sum((x - mean(x))^2)/(length(x))
}

pop_sd <- function(x) {
    sqrt(pop_v(x))
}
pop_sd(x)
```

###weight 

 mean
```{r}
population_mean_weight <- mean(zombies$weight)
population_mean_weight
```

standard deviation 
```{r}
x <- zombies$weight
pop_sd(x)
```

###age

 mean
```{r}
population_mean_age <- mean(zombies$age)
population_mean_age
```

standard deviation 
```{r}
x <- zombies$age
pop_sd(x)
```

###number of zombies killed

 mean
```{r}
population_mean_zombies_killed <- mean(zombies$zombies_killed)
population_mean_zombies_killed
```

standard deviation 
```{r}
x <- zombies$zombies_killed
pop_sd(x)
```

###years of education 

 mean
```{r}
population_mean_years_of_education <- mean(zombies$years_of_education)
population_mean_years_of_education
```

standard deviation 
```{r}
x <- zombies$years_of_education
pop_sd(x)
```

NK: I found that if I selected the quantitative variables from the data frame, I was then able to use cbind and apply to quickly calculate all of the means and standard deviations and produce a new data frame with this information. I specified the new columns as needed, and included the calculations as the definition of each column. This saved me a lot of time, although it might defeat the learning opportunity involved in walking step by step through each calculation. 

##2. Use {ggplot} to make boxplots of each of these variables by gender.

```{r}
library(ggplot2)
zom.height <- zombies$height
```

###height 

```{r}
z <- ggplot(data = zombies, aes(x = gender, y = height))
z <- z + geom_boxplot()

z
```

###weight

```{r}
z <- ggplot(data = zombies, aes(x = gender, y = weight))
z <- z + geom_boxplot()

z
```

###age

```{r}
z <- ggplot(data = zombies, aes(x = gender, y = age))
z <- z + geom_boxplot()

z
```

###number of zombies killed

```{r}
z <- ggplot(data = zombies, aes(x = gender, y = zombies_killed))
z <- z + geom_boxplot()

z
```

###years of education

```{r}
z <- ggplot(data = zombies, aes(x = gender, y = years_of_education))
z <- z + geom_boxplot()

z
```

##3. Use {ggplot} to make scatterplots of height and weight in relation to age. Do these variables seem to be related? In what way?

###height 
```{r}

z <- ggplot(data = zombies, aes(x = age, y = height, color = factor(gender)))
z <- z + geom_point()
z <- z + theme(legend.position = "bottom", legend.title = element_blank())  
z
```

###weight 
```{r}
z <- ggplot(data = zombies, aes(x = age, y = weight, color = factor(gender)))
z <- z + geom_point()
z <- z + theme(legend.position = "bottom", legend.title = element_blank())  
z
```
both the height and the weight plots show a direct, positive relationship between age and height, and age and weight. I found it surprising that there seems to be a stronger relationship between age and height than there is for age and weight... I would expect there to be a stronger relationship between age and height in people's 20's because humans don't usually grow taller after like 18.... or maybe thats just me... 

NK: We have similar solutions here!

##4. Using histograms and Q-Q plots, check whether the quantitative variables seem to be drawn from a normal distribution. Which seem to be and which do not (hint: not all are drawn from the normal distribution)? For those that are not normal, can you determine from which common distribution they are drawn?

###height 

```{r}
h <- zombies$height
hist(h) #histogram shows a normal distribution around the mean 
population_mean_height
```

```{r}
qqnorm(h, main = "QQ plot of Zombie Killer Height")
qqline(h, col = "gray") #normally distributed 
```

###weight

```{r}
w <- zombies$weight
hist(w) #histogram shows a normal distribution around the mean 
population_mean_weight
```

```{r}
qqnorm(w, main = "QQ plot of Zombie Killer Weight")
qqline(w, col = "gray") #normally distributed 
```
###age 

```{r}
a <- zombies$age
hist(a) #histogram shows a normal distribution around the mean 
population_mean_age
```

```{r}
qqnorm(a, main = "QQ plot of Zombie Killer Age")
qqline(a, col = "gray") #normally distributed 
```

###number of zombies killed 

```{r}
k <- zombies$zombies_killed
hist(k) #histogram is skewed towards lower numbers 
population_mean_zombies_killed
```

```{r}
qqnorm(k, main = "QQ plot of Zombie Kills")
qqline(k, col = "gray") #WOAH!! super weird! 
```

###years of education 

```{r}
e <- zombies$years_of_education
hist(e) #histogram is skewed towards lower numbers 
population_mean_years_of_education
```

```{r}
qqnorm(e, main = "QQ plot of Zombie Killer Years of Education")
qqline(e, col = "gray") #QQplot looks just like plot zombie kills 
```

##For those that are not normal, can you determine from which common distribution they are drawn? 

number of zombies killed and years of education are not normal because these data are not random. To determine from which common distribution they are drawn, you can look at the data itself...
```{r}
zombies$zombies_killed
zombies$years_of_education
```
these data are independent counts of integer numbers, which would follow the Poisson distribution 

NK: I like that you immediately comment your observations for each plot next to the code. I will do that next time as well. I used ggplot, but it looks like we came to the same conclusions, so I don't think it matters which one you use to visualize the data. Ahhh I forgot to answer the question about which distribution is used for the non-normal data, but your explanation makes sense! 

##5. Now use the sample() function to sample ONE subset of 30 zombie survivors (without replacement) from this population and calculate the mean and sample standard deviation for each variable.

###height 
```{r}
?sample()
zombies
sample_height <- sample(zombies$height, size = 30, replace = FALSE, prob = NULL) 
sample_height
mean(sample_height)
sd(sample_height)
```

##Also estimate the standard error for each variable, and construct the 95% confidence interval for each mean. Note that for the variables that are not drawn from the normal distribution, you may need to base your estimate of the CIs on slightly different code than for the normal…

NK: it's nice that you included ?sample at the beginning of the chunk to quickly explain its use. 

standard error

```{r}
se_height <- (sd(sample_height)/sqrt(30))
se_height
```

construct 95% confidence interval for the mean 

```{r}
x <- sample_height
m <- mean(sample_height)

normalCI = function(x, CIlevel = 0.95) {
    upper = m + qnorm(1 - (1 - CIlevel)/2) * sqrt(var(x)/length(x))
    lower = m + qnorm((1 - CIlevel)/2) * sqrt(var(x)/length(x))
    ci <- c(lower, upper)
    return(ci)
}
normalCI(x, 0.95)
```

###weight 

  create sample  
```{r}
sample_weight <- sample(zombies$weight, size = 30, replace = FALSE, prob = NULL) 
sample_weight
```

  mean 
```{r}
mean(sample_weight)
```
  standard deviation 
```{r}
sd(sample_weight)
```

  standard error 
```{r}
se_weight <- (sd(sample_weight)/sqrt(30))
se_weight
```
  95% Confidence Interval 
```{r}
x <- sample_weight
m <- mean(sample_weight)
normalCI(x, 0.95)
```

###age

  create sample  
```{r}
sample_age <- sample(zombies$age, size = 30, replace = FALSE, prob = NULL) 
sample_age
```

  mean 
```{r}
mean(sample_age)
```
  standard deviation 
```{r}
sd(sample_age)
```

  standard error 
```{r}
se_age <- (sd(sample_age)/sqrt(30))
se_age
```
  95% Confidence Interval 
```{r}
x <- sample_age
m <- mean(sample_age)
normalCI(x, 0.95)
```

###number of zombies killed

  create sample  
```{r}
sample_zombies_killed <- sample(zombies$zombies_killed, size = 30, replace = FALSE, prob = NULL) 
sample_zombies_killed
```
  mean 
```{r}
mean(sample_zombies_killed)
```

  standard deviation 
```{r}
sd(sample_zombies_killed)
```

  standard error 
```{r}
se_zombies_killed <- (sd(sample_zombies_killed)/sqrt(30))
se_zombies_killed
```

95% Confidence Interval 

Write function to calculate confidence intervals under Poisson Distribution 
```{r}
x <- mean(sample_zombies_killed)


poisson95CI = function(x) {
    upper = x + 1.96*(sqrt(x/30))
    lower = x - 1.96*(sqrt(x/30))
    ci <- c(lower, upper)
    return(ci)
}

poisson95CI(x)
```

###years of education 
  create sample  
```{r}
sample_years_of_education <- sample(zombies$years_of_education, size = 30, replace = FALSE, prob = NULL) 
sample_years_of_education
```
  mean 
```{r}
mean(sample_years_of_education)
```

  standard deviation 
```{r}
sd(sample_years_of_education)
```

  standard error 
```{r}
se_years_of_education <- (sd(sample_years_of_education)/sqrt(30))
se_years_of_education
```

95% Confidence Interval 
```{r}
x <- mean(sample_years_of_education)
poisson95CI(x)
```

NK: I can't tell how similar our numbers would be, but I see you used poisson95CI here. I would think this would work, given you identified the poisson distribution earlier. I ended up using a for loop to bootstrap my confidence intervals for the non-normally distributed data. 

##6. Now draw 99 more random samples of 30 zombie apocalypse survivors, and calculate the mean for each variable for each of these samples. Together with the first sample you drew, you now have a set of 100 means for each variable (each based on 30 observations), which constitutes a sampling distribution for each variable. 

###height 
```{r}
set <- NULL
n <- zombies$height
for (i in 1:99) {
    set[i] <- mean(sample(n, size = 30, replace = FALSE, prob = NULL))
}
sampling_distribution <- set[]
sampling_distribution

```

##What are the means and standard deviations of this distribution of means for each variable? How do the standard deviations of means compare to the standard errors estimated in [5]? What do these sampling distributions look like (a graph might help here)? Are they normally distributed? What about for those variables that you concluded were not originally drawn from a normal distribution?

```{r}
mean(sampling_distribution)
sd(sampling_distribution)
```

This standard deviation of the sampling distribution is actually very similar to the standard error from [5] - 0.7010483, COOL! 

```{r}
hist(sampling_distribution) #looks like a normal distrubution 
```
###weight 
```{r}
set <- NULL
n <- zombies$weight
for (i in 1:99) {
    set[i] <- mean(sample(n, size = 30, replace = FALSE, prob = NULL))
}
sampling_distribution <- set[]
sampling_distribution

```

```{r}
mean(sampling_distribution)
sd(sampling_distribution)
```
this is similar to the standard error value fron 5 (3.855582)

```{r}
hist(sampling_distribution) #looks like a normal distribution 
```

###age
```{r}
set <- NULL
n <- zombies$age
for (i in 1:99) {
    set[i] <- mean(sample(n, size = 30, replace = FALSE, prob = NULL))
}
sampling_distribution <- set[]
sampling_distribution

```

```{r}
mean(sampling_distribution)
sd(sampling_distribution)
```
this is similar to the standard error value from 5 (0.4423582)

```{r}
hist(sampling_distribution) #looks like a normal distribution 
```

###zombies killed

```{r}
set <- NULL
n <- zombies$zombies_killed
for (i in 1:99) {
    set[i] <- mean(sample(n, size = 30, replace = FALSE, prob = NULL))
}
sampling_distribution <- set[]
sampling_distribution

```

```{r}
mean(sampling_distribution)
sd(sampling_distribution)
```
this is similar to the standard error value from 5 (0.3571385)

```{r}
hist(sampling_distribution) #looks like a normal distribution 
```

###years of education 
```{r}
set <- NULL
n <- zombies$years_of_education
for (i in 1:99) {
    set[i] <- mean(sample(n, size = 30, replace = FALSE, prob = NULL))
}
sampling_distribution <- set[]
sampling_distribution

```

```{r}
mean(sampling_distribution)
sd(sampling_distribution)
```
this is similar to the standard error value from 5 (0.3956178)

```{r}
hist(sampling_distribution) #looks like a normal distribution 
```

NK: We have similar answers here. I really wanted to somehow run the data all in one or two goes rather than one column at a time, but I couldn't figure out how. 

Challenges: 

5. Error in sample.int(length(x), size, replace, prob) : cannot take a sample larger than the population when 'replace = FALSE'

original code: sample(zombies, size = 30, replace = FALSE, prob = NULL), needed to specify zombies$variable 
